---
layout: page
title: hnRNP A2 interactome - Fig3
---

```{r options, echo=FALSE}
library(knitr)
library(DESeq2)
library(ggplot2)
library(clusterProfiler)
library(ggplot2)
library(biomaRt)
library(sva)
library(org.Rn.eg.db)
library(dplyr)
library(ggrepel)
```

## Directory of reads

```{r messages=FALSE}
directory <- file.path("rip/counts_stranded/gene/")
```


#Ensembl Annotation

```{r cars}
if(exists("ensembl")==F){
ensembl<-useMart("ensembl")}
if(exists("rat")==F){
rat<-useDataset("rnorvegicus_gene_ensembl", mart=ensembl)}
listAttributes(rat)
rat_genes <- getBM(attributes=c('external_gene_name',"ensembl_gene_id","gene_biotype"), mart = rat)
rat_genes_go <- getBM(attributes=c('external_gene_name',"ensembl_gene_id", "ensembl_transcript_id","go_id","name_1006", "namespace_1003"), mart = rat)
for(i in 1:length(rat_genes[,1])){
  if(rat_genes[i,1]==""){
    rat_genes[i,1]<-rat_genes[i,2]
  }
}
for(i in 1:length(rat_genes_go[,1])){
  if(rat_genes_go[i,1]==""){
    rat_genes_go[i,1]<-rat_genes_go[i,2]
  }
}
```



############################################## Interactome hnRNP A2 - Pipeline

# Sample Table

Depending on the interactome desired, run one or other sample table. 
IgG and IgGAb are the same files, as they are a pool of treated and untreated mock precipitation (IgG), but duplicated for easier reproducibility in this script.

```{r}
#Sample Table for Control Condition

sampleFiles <- list.files(directory)
sampleName<-sapply(sampleFiles %>% str_split("_"), "[[", 1)  
sampleName <- sampleName[-grep("Ab", sampleName)]
sampleExperiment<-sampleName %>% str_extract("Input|RIP|IgG")
sampleRIP<-sampleExperiment
sampleRIP[grep("RIP|IgG",sampleRIP)]<-"yes"
sampleRIP[grep("Input",sampleRIP)]<-"no"
samplePair <- sampleName %>% str_extract('[0-9]')
sampleIndividual <- sampleName %>% str_replace("InputC|RIPC|IgG",'') 
sampleFiles <- sampleFiles[-grep("Ab", sampleFiles)]
sampleTable <- data.frame(sampleName = sampleName,
                          sampleFiles = sampleFiles,
                          Experiment = sampleExperiment,
                          RIP=sampleRIP,
                          Individual = sampleIndividual,
                          Pair = samplePair)
sampleTable$Pair <- factor(sampleTable$Pair)
sampleTable$Experiment <- factor(sampleTable$Experiment, levels = c("Input","IgG", "RIP"))
sampleTable$RIP <- factor(sampleTable$RIP, levels = c("no", "yes"))
sampleTable$Individual <- factor(sampleTable$Individual)
rownames(sampleTable)<-sampleTable$sampleFiles
```


```{r}
#Sample Table for Abeta Condition

sampleFiles <- list.files(directory)
sampleName<-sapply(sampleFiles %>% str_split("_"), "[[", 1)  
sampleName <- sampleName[-grep("C", sampleName)]
sampleExperiment<-sampleName %>% str_extract("Input|RIP|IgG")
sampleRIP<-sampleExperiment
sampleRIP[grep("RIP|IgG",sampleRIP)]<-"yes"
sampleRIP[grep("Input",sampleRIP)]<-"no"
samplePair <- sampleName %>% str_extract('[0-9]')
sampleIndividual <- sampleName %>% str_replace("InputC|RIPC|IgG",'') 
sampleFiles <- sampleFiles[-grep("Ab", sampleFiles)]
sampleTable <- data.frame(sampleName = sampleName,
                          sampleFiles = sampleFiles,
                          Experiment = sampleExperiment,
                          RIP=sampleRIP,
                          Individual = sampleIndividual,
                          Pair = samplePair)
sampleTable$Pair <- factor(sampleTable$Pair)
sampleTable$Experiment <- factor(sampleTable$Experiment, levels = c("Input","IgG", "RIP"))
sampleTable$RIP <- factor(sampleTable$RIP, levels = c("no", "yes"))
sampleTable$Individual <- factor(sampleTable$Individual)
rownames(sampleTable)<-sampleTable$sampleFiles
```

# Import Data to DESeq2


```{r}
dds <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                       directory = directory,
                                       design= ~ Pair + Experiment)
rowData(dds)<-rat_genes[match(rownames(dds), rat_genes$ensembl_gene_id),]
```


#Filter by counts

Expressed > 5 at least in two samples of the three replicates per condition. 

```{r}
dds_keep<-dds
keep <- "Values"
# unevaluated chunk...
for(i in 1:3){
  j<-i*3
  h<-j-2
  keep_p <- rownames(dds_keep)[rowSums(counts(dds_keep)[,c(h:j)]>= 2)>= 2] 
  keep <- c(keep, keep_p)
}
keep <- keep[-1]
keep <- unique(keep)
dds_keep<-dds[keep,]
```

# Processing DESeq2

```{r}
dds_keep <- estimateSizeFactors(dds_keep)
dds_keep <- estimateDispersions(dds_keep)
dds_keep <- nbinomWaldTest(dds_keep)
```


# Normalization

```{r}
rld_keep <- rlog(dds_keep, blind=FALSE)
```


# Results

```{r}
dds_keep<-DESeq(dds_keep)
res_rip<-results(dds_keep, contrast = list("Experiment_RIP_vs_Input","Experiment_IgG_vs_Input"), alpha = 0.05)
```

# Visualization

```{r}
ma_plot_data <- as.data.frame(res_rip)
ma_plot_data$padj[is.na(ma_plot_data$padj)]<-1
n<-length(which(ma_plot_data$padj < 0.05 &  ma_plot_data$log2FoldChange > 0))
ggplot(data = ma_plot_data, aes(x = baseMean, y = log2FoldChange)) +
  geom_point(color = ifelse(ma_plot_data$padj < 0.05 &  ma_plot_data$log2FoldChange > 0, "steelblue4", "grey20"), alpha = 0.5)+scale_x_log10()+xlab("Log10Mean of Normalized counts") + annotate(geom="label", x=15, y=7.5, label=paste("◉ ↑ RIP-seq", n, collapse = " "),
              color="steelblue4")
```


# Gene Ontology Enrichment


```{r}
ego_ab <- enrichGO(gene = results %>% filter(padj < 0.05) %>% pull(external_gene_name),
                keyType = "ENSEMBL",
                OrgDb = org.Rn.eg.db, 
                ont = "ALL", 
                pAdjustMethod = "BH", 
                qvalueCutoff = 0.05, 
                readable = TRUE)
```



